### 2013-06-14 doc/pgas13/relatedwork.tex
Rework related work based on Simon's feedback.

Flat combining~\cite{flatCombining} is the latest of many techniques for allowing a single thread to perform the updates of others. Previous techniques performed combining hierarchically using fixed trees~\cite{yew:combining-trees}, dynamic trees (or ``funnels'')~\cite{funnels}, and randomized trees~\cite{edtrees}. However, the observation of the flat combining work was that in practice, a cleverly-implemented \emph{flat} publication list can outperform more complicated schemes.
Follow-on work introduced parallel flat-combining~\cite{scalableFCQueues}, in which multiple publication lists are combined in parallel and their remainders serialized.
Flat combining was also applied in the NUMA (non-uniform memory access) domain with scalable hierarchical locks~\cite{fcNUMALocks}, which improves on prior work on hierarchical locking by leveraging flat-combining's publication mechanism which amortizes the cost of synchronization.

Our work extends the flat-combining paradigm even further to the PGAS domain, where only software provides the illusion of a physically-distributed memory. The relatively higher cost of communication in PGAS makes flat combining even more compelling. In addition, the physical isolation per node combined with Grappa's engineered per-core isolation guarantees, introduce a novel opportunity to apply combining, which led to our ``combining proxy'' design.

>>>>

Many techniques performed combining hierarchically to reduce synchronization and hot-spotting, differing mainly in the structures used to do combining: fixed trees~\cite{yew:combining-trees}, dynamic trees (or ``funnels'')~\cite{funnels,MAMA}, and randomized trees~\cite{edtrees}. In particular, MAMA~\cite{MAMA}, built for the MTA-2, leveraged funnels to better utilize hardware optimized for massive concurrency.
On the other hand, \emph{flat} combining~\cite{flatCombining} observed that in multicore systems, hierarchical schemes have too much overhead, and a cleverly-implemented \emph{flat} publication list can perform better.
Follow-on work introduced parallel flat-combining~\cite{scalableFCQueues}, in which multiple publication lists are combined in parallel and their remainders serialized.
Flat combining was also applied in the NUMA (non-uniform memory access) domain with scalable hierarchical locks~\cite{fcNUMALocks}, which improves on prior work on hierarchical locking by leveraging flat-combining's publication mechanism which amortizes the cost of synchronization.

Our work extends the flat-combining paradigm even further to the PGAS domain, where only software provides the illusion of a physically-distributed memory. The relatively higher cost of communication in PGAS makes flat combining even more compelling. Within-core combining is naturally done in a \emph{flat} way. In addition, the physical isolation per node combined with Grappa's engineered per-core isolation guarantees, create a novel situation for combining, which led to our \emph{combining proxy} design.
